Hydras \& Co.

### 2022-03-02 doc/part-hydras.tex
Small changes in gaia-chapter; add an ON instance for Mathcomp finite ordinals
corrections in gaia-chapter
gaia-hydras first draft
Remove ocons (#124)
* minor script simplifications

* minor script simplifications

* make T1's constructors gaia-compatible

@@

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
level 2/.style={sibling distance=35mm},  
level 3/.style={sibling distance=17mm}]
\node  {ocons}
  child {  node {ocons}
            child { node {ocons} child {node {zero}} child {node{0}} child{node{zero}}}
         child {node {0}}
         child {node {zero}}}
    child {node {0}}
   child {node {ocons} 
 child { node {ocons} child {node {zero}} child {node{2}} child{node{zero}}}
  child {node {4}}
         child {node {ocons} child {node {zero}} child {node{1}} child{node{zero}}}};

@@

\paragraph{Remark}
For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{ocons $\alpha$ $n$ $\beta$}) is intended to represent the
ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
definition. But this replacement would take a lot of time \dots{}

@@

Unfortunately, it won't work. Let us consider some well-formed term
 $\alpha=\texttt{ocons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
 through \texttt{LT}. For proving the accessibility of $\alpha$, we have to consider
any well formed term \(\delta\) such that \(\delta<\alpha\). 
But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).

@@

The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
\(\gamma\);
for instance \(\delta\) may be of the form
\texttt{ocons $\beta$ $p'$  $\gamma'$},
where    \(p' < n\).
Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.

@@

First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
strongly accessible too (\emph{i.e.} any well formed
term (\texttt{ocons $\alpha$ $n$ $\beta$})  is accessible).

@@

Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{ocons alpha n beta}) is in normal form. 
We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.

@@

    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
    \label{sec:orgheadline87}
    \label{hydra-variant}

@@

    % \begin{Coqsrc}
    %  Lemma oplus_rect:
    %       forall P: T1 -> T1 -> T1 -> Type, 
    %         (forall a:T1, P zero a a) ->
    %         (forall a: T1, P a zero a) ->
    %         (forall a1 n1 b1 a2 n2 b2 o,
    %            compare a1 a2 = Gt ->
    %            P b1 (ocons a2 n2 b2) o ->
    %            P (ocons a1 n1 b1) (ocons a2 n2 b2)
    %              (ocons a1 n1 o)) ->
    %         (forall a1 n1 b1 a2 n2 b2 o,
    %            compare a1 a2 = Lt ->
    %            P (ocons a1 n1 b1) b2 o ->
    %            P (ocons a1 n1 b1) (ocons a2 n2 b2) 
    %            (ocons a2 n2 o)) ->
    %         (forall a1 n1 b1 a2 n2 b2 o,
    %            compare a1 a2 = Eq ->
    %            P b1 b2 o ->
    %           P (ocons a1 n1 b1) (ocons a2 n2 b2)
    %             (ocons a1 (S (n1 + n2)%nat) o)) ->
    %          forall a b, P a b (oplus a b).
    % Proof with auto.
    %    induction a.
    %    -    intro; simpl; destruct b;auto.
    %    -   induction b.
    %        + apply X0.
    %        + case_eq (compare a1 b1).
    %          * intro Comp; unfold oplus; rewrite Comp.
    %            cbn; apply X3 ...
    %          * intro Comp; cbn; rewrite Comp; apply X2...
    %          * intro Comp; cbn; rewrite Comp ...
    %  Defined.

@@

>>>>

The introduction of Jos\'e Grimm's report~\cite{grimm:hal-00911710} contains a nice presentation of the main properties of ordinals.

@@

\begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.5, level 1/.style={sibling distance=6cm},
level 2/.style={sibling distance=35mm},  
level 3/.style={sibling distance=17mm}]
\node  {cons}
  child {  node {cons}
            child { node {cons} child {node {zero}} child {node{0}} child{node{zero}}}
         child {node {0}}
         child {node {zero}}}
    child {node {0}}
   child {node {cons} 
 child { node {cons} child {node {zero}} child {node{2}} child{node{zero}}}
  child {node {4}}
         child {node {cons} child {node {zero}} child {node{1}} child{node{zero}}}};

@@

\paragraph{Remark}
For simplicity's sake, we chose to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the construction (\texttt{cons $\alpha$ $n$ $\beta$}) is intended to represent the
ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
definition. But this replacement would take a lot of time \dots{}

@@

Unfortunately, it won't work. Let us consider some well-formed term
 $\alpha=\texttt{cons $\beta\;n\;\gamma$}$, and assume that \(\beta\) and \(\gamma\) are accessible
 through \texttt{LT}. For proving the accessibility of $\alpha$, we have to consider
any well formed term \(\delta\) such that \(\delta<\alpha\). 
But nothing guarantees that \(\delta\)  is strictly  less than \(\beta\) nor \(\gamma\), and we cannot use the induction hypotheses on   \(\beta\) nor \(\gamma\).

@@

The problem comes from the hypothesis \texttt{Hdelta}. It does not prevent  \(\delta\) to be bigger that \(\beta\) or
\(\gamma\);
for instance \(\delta\) may be of the form
\texttt{cons $\beta$ $p'$  $\gamma'$},
where    \(p' < n\).
Thus, the induction hypotheses \texttt{IHbeta} and \texttt{IHgamma}  are useless for finishing our proof.

@@

First, we prove that, for  any \texttt{LT}-accessible term $\alpha$, $\alpha$ is 
strongly accessible too (\emph{i.e.} any well formed
term (\texttt{cons $\alpha$ $n$ $\beta$})  is accessible).

@@

Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{cons alpha n beta}) is in normal form. 
We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
and natural induction on \texttt{n}, that (\texttt{cons alpha n beta}) is accessible.
The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.

@@

    \subsection{Natural sum (a.k.a. Hessenberg's  sum)}
    \label{sec:orgheadline87}
    \label{hydra-variant}
    \label{sect:hessenberg-def}

@@

    % \begin{Coqsrc}
    %  Lemma oplus_rect:
    %       forall P: T1 -> T1 -> T1 -> Type, 
    %         (forall a:T1, P zero a a) ->
    %         (forall a: T1, P a zero a) ->
    %         (forall a1 n1 b1 a2 n2 b2 o,
    %            compare a1 a2 = Gt ->
    %            P b1 (cons a2 n2 b2) o ->
    %            P (cons a1 n1 b1) (cons a2 n2 b2)
    %              (cons a1 n1 o)) ->
    %         (forall a1 n1 b1 a2 n2 b2 o,
    %            compare a1 a2 = Lt ->
    %            P (cons a1 n1 b1) b2 o ->
    %            P (cons a1 n1 b1) (cons a2 n2 b2) 
    %            (cons a2 n2 o)) ->
    %         (forall a1 n1 b1 a2 n2 b2 o,
    %            compare a1 a2 = Eq ->
    %            P b1 b2 o ->
    %           P (cons a1 n1 b1) (cons a2 n2 b2)
    %             (cons a1 (S (n1 + n2)%nat) o)) ->
    %          forall a b, P a b (oplus a b).
    % Proof with auto.
    %    induction a.
    %    -    intro; simpl; destruct b;auto.
    %    -   induction b.
    %        + apply X0.
    %        + case_eq (compare a1 b1).
    %          * intro Comp; unfold oplus; rewrite Comp.
    %            cbn; apply X3 ...
    %          * intro Comp; cbn; rewrite Comp; apply X2...
    %          * intro Comp; cbn; rewrite Comp ...
    %  Defined.

@@

% ---------------------------------------------------
\input{gaia-chapter}
